Density dependence

### 2020-09-17 R/density dependence.Rmd
figuring out crossbasis smooth

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
```

@@

check_overdispersion(mpois2) #actually underdispersed

@@

```{r}
fixef(m_growth) %>% exp()
```

@@

```{r}
Anova(m_surv2)
fixef(m_surv2) %>% plogis()
```

Marginal, and only immediate grid cell.  Density has positive affect on survival.  So, probably just good patches.

@@

```{r}
Anova(m_flwr)
fixef(m_flwr) %>% plogis()
```

@@

@@

- Check number of inflorescencesa given flowered?
- ~~check that survival is coded right so that it is a function of `ht_prev`~~
- try negative binomial distribution for shoot number
- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- ~~incorporate 1-ha fragments and use habitat type as a predictor.~~
- Implement FLM approach to density dependence.

>>>>

conflict_prefer("lag", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

@@

check_overdispersion(mpois) #actually underdispersed

@@

```{r}
fixef(m_growth)
```

@@

```{r}
Anova(m_surv2)
fixef(m_surv2)
```

Marginal effects on survival.

@@

```{r}
Anova(m_flwr)
fixef(m_flwr)
```

@@

Probability of flowering increases with size and is lower in continuous forest

# Kernel approach

So it seems like there are effect of density at least out to 20m.  I'd like to take the approach in Teller et al. 2016, but the code is so convoluted that it is not adaptable. I want to start by trying to adapt the `dlnm` approach instead.

## Set up the crossbasis

```{r}
Q <-
  ha_sub %>%
  select(density_0_10:density_50_60) %>% 
  as.matrix()

L <-
  matrix(seq(5, 55, length.out = 6),
         nrow = nrow(Q), 
         ncol = 6,
         byrow = TRUE)
```

## GAM

```{r}
library(dlnm)
library(mgcv)

# Q_log <- log(Q+1)
g_growth <-
  gam(shts_next ~ s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          k = c(5, 4),
          xt = list(bs = "cr")),
      family = gaussian(link = "log"),
      data = ha_sub,
      method = "REML")
```

## Predict

```{r}
x <- crosspred("Q", model = g_growth, cumul = TRUE)
summary(x)
```

## Plot crossbasis

### Contour plot

```{r}
plot.crosspred(x,
               ptype = "contour",
               xlab = "#shoots/100m^2",
               ylab = "distance (m)",
               cumul = FALSE)
```
So this shows negative density dependence that seems to be strongest at about 40-50 m away from the focal plant.

### Slices

We can also plot slices through the contour plot at different distances or different shoot densities

```{r}
plot.crosspred(x, ptype = "slice", var = 100, xlab = "distance (m)", main = "Effect of 100 shoots/100m^2")
plot.crosspred(x, ptype = "slice", var = 10, xlab = "distance (m)", main = "Effect of 10 shoots/100m^2")
```

### Cumulative

I think this shows the cumulative effect of density summed/integrated over all distances.  So, for example, a plant with about 30 shoots/100m^2 surrounding it out to the maximum distance (60m) would have the largest number of shoots, on average.

```{r}
plot(x, ptype = "overall", xlab = "density (shoots/100m^2)", cumul = TRUE)
```

There is no built-in function to get cumulative effects of density over all distances.

## Other shapes

It is odd that the strength of competition seems to be increasing with distance.  I'd like to be able to force the distance dimension to be monotonically decreasing as was done in Teller et al (I think?).  Teller does this *after* fitting a spline using `mk_dist_wts()`, which is a function that takes the spline coefficients, and predicted values from the spline, and transforms it into a non-negative, monotonic decreasing function.

You can override one of the dimensions of the cross basis with any custom function.

```{r}
#use arglag to override lag dimension of crossbasis, use argvar to override plant density dimension.
lin <- dlnm:::lin
xt <- list(bs = "cr", arglag = list(fun = "lin"))

g_growth <- 
  gam(shts_next ~ s(log_shts) +
        habitat +
        s(Q, L,
          bs = "cb",
          k = c(5, 4),
          fx = c(F, T), #turns off penalization for the lag dimension
          xt = xt),
      family = gaussian(link = "log"),
      data = ha_sub,
      gamma = 1.5)

gam.check(g_growth)
```

```{r}
x <- crosspred("Q", model = g_growth)
plot.crosspred(x, ptype = "contour", xlab = "#shoots/100m^2", ylab = "distance (m)")

plot.crosspred(x, ptype = "slice", var = 100)
plot.crosspred(x, ptype = "slice", var = 10)
```
Now the kernel is the same shape at every value of density.  However, this relationship doesn't make any sense.  Fewer knots?  More penalty?  Custom function?

```{r}
anova(g_growth)
```

@@

- ~~is density at 0 distance including the focal plant?  It shouldn't.~~
- Check number of inflorescences given flowered?
- ~~log-transform density (nah, not right now.  DOesn't help with convergence)~~
- ~~check that survival is coded right so that it is a function of `ht_prev`~~
- try negative binomial distribution for shoot number
- ~~Density should be total shoot number~~
- ~~Should actually be density in previous year~~
- ~~Shoots should be poisson distribution, not log-transformed?~~
- ~~incorporate 1-ha fragments and use habitat type as a predictor.~~
- Implement FLM approach to density dependence.
